\texttt{JAX-COSMO}: An End-to-End Differentiable and GPU Accelerated Cosmology Library

### 2023-04-19 paper_OJA.tex
minor update

The angular power spectra $C_\ell^{ij}$ for the probes $(i,j)$ and for redshift bin window selections are computed in the first order Limber approximation \citep{PhysRevD.78.123506}:
\begin{align}
    C_\ell^{i,j} \approx \left(\ell+\frac{1}{2}\right)^{m_i+m_j}\int\frac{\mathrm{d}a}{c^2\chi^2}\frac{\mathrm{d}\chi}{\mathrm{d}a}K_i(\chi)K_j(\chi)\,P\left(k=\frac{\ell+1/2}{\chi},z\right),\label{eq:Cell_limber}
\end{align}
The $m_i$ factors are $(0,-2)$ for the galaxy clustering and weak lensing, respectively, and each $K(z)$ function represents a single tomographic redshift bin's number density \resubnote{in $z=(1/a)-1$. The integration lower bound $a_{ij}$ is defined from the maximum redshift involved  and  Simpson's rule is used  with a regular grid of 512 knots. The $K(z)$} tracers are implemented as two kernel functions:

@@

\resubnote{Despite this high precision with respect to the CCL library, the current \jaxcosmo\ implementation will have to be reviewed to tackle Stage IV accuracy requirements. For instance the non-limber $C_\ell$  computation \cite{angpow, n5k} would be a limitation needing revision in a future release.}

@@

A second approach would be to develop emulators of a fully-fledged Boltzmann code. Emulators based on neural networks or Gaussian processes are themselves automatically differentiable with respect to cosmological parameters. In fact, the literature is now rich in examples of such emulators \citep[e.g.][and references therein]{Gunther_2022, nygaard,cosmopower,cosmicnet, emucmb, bacco1, bacco2}. After validating their accuracy against a reference CAMB or CLASS implementation, they could be directly integrated as a plug-and-play replacement for the computation of the matter power spectrum. At this time, it seems that using emulators will be the most straightforward approach to bring more accurate models to \jaxcosmo. We believe, though,  that one of the reason for the wide diversity in this is a lack of standardization - a unified interface and validation suite for such methods would provide a much simpler comparison between them and enable wider usage.

>>>>

The angular power spectra $C_\ell^{ij}$ for the probes $(i,j)$ and for redshift bin window selections are computed in the first order Limber approximation \citep{PhysRevD.78.123506}:
\begin{align}
    C_\ell^{i,j} \approx \left(\ell+\frac{1}{2}\right)^{m_i+m_j}\int_{a_{i,j}}^1\frac{\mathrm{d}a}{c^2\chi^2}\frac{\mathrm{d}\chi}{\mathrm{d}a}K_i(\chi)K_j(\chi)\,P\left(k=\frac{\ell+1/2}{\chi},z\right),\label{eq:Cell_limber}
\end{align}
The $m_i$ factors are $(0,-2)$ for the galaxy clustering and weak lensing, respectively, and each $K(z)$ function represents a single tomographic redshift bin's number density \resubnote{in $z=(1/a)-1$. The integration lower bound $a_{ij}$ is defined from the maximum redshift involved  and  Simpson's rule is used  with a regular grid of 512 knots. The $K(z)$} tracers are implemented as two kernel functions:

@@

\resubnote{Despite this high precision with respect to the CCL library, the current \jaxcosmo\ implementation will have to be reviewed to tackle Stage IV accuracy requirements. For instance the non-limber $C_\ell$  computation \citep{n5k, angpow} would be a limitation needing revision in a future release.}

@@

A second approach would be to develop emulators of a fully-fledged Boltzmann code. Emulators based on neural networks or Gaussian processes are themselves automatically differentiable with respect to cosmological parameters. In fact, the literature is now rich in examples of such emulators \citep[e.g.][and references therein]{bacco1, bacco2, Gunther_2022, nygaard,cosmopower,cosmicnet, emucmb}. After validating their accuracy against a reference CAMB or CLASS implementation, they could be directly integrated as a plug-and-play replacement for the computation of the matter power spectrum. At this time, it seems that using emulators will be the most straightforward approach to bring more accurate models to \jaxcosmo. We believe, though,  that one of the reason for the wide diversity in this is a lack of standardization - a unified interface and validation suite for such methods would provide a much simpler comparison between them and enable wider usage.
